An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-18 draft.tex
More description and images

@@

\item The Controller process is responsible for keeping track of the progress of 
the search. It pulls current subpopulations from the Output queue, inspects the metadata 
and if an optimal solution has been found or the maximum number of iteration has been
reached it signals the stop of the execution. Otherwise, it passes the stream of messages 
to a migration process, where subpopulations are mixed with each other. New populations 
are generated from this migration, and they are again pushed to the Input queue to continue 
in a loop. The Controller is also responsible of logging the metadata recived with the
messages.

\end{enumerate}

Based on a reactive architecture \cite{guervos2018introducing} , each
subpopulation is represented as an event message that is pushed asynchronously
into a message queue. Incoming messages trigger stateless functions that receive
the subpopulation and proceed to run an algorithm, using the parameters and
population included in the message. After the specified number of iterations,
each stateless function returns the evolved subpopulation by again pushing a
message to another queue, used for receiving the resulting subpopulations.
Subpopulations are received from the queue by a controller that is responsible
for mixing the individuals from different subpopulations and producing new
subpopulations. These new subpopulations are pushed again by the controller into
the message queue, creating a loop. The cycle stops when the controller receives
a subpopulation containing a candidate solution that satisfies a particular
condition, or a maximum number of messages were received.

@@

\subsection{Controller} 
\label{edm}

>>>>

% Based on a reactive architecture \cite{guervos2018introducing} , each
% subpopulation is represented as an event message that is pushed asynchronously
% into a message queue. Incoming messages trigger stateless functions that receive
% the subpopulation and proceed to run an algorithm, using the parameters and
% population included in the message. After the specified number of iterations,
% each stateless function returns the evolved subpopulation by again pushing a
% message to another queue, used for receiving the resulting subpopulations.
% Subpopulations are received from the queue by a controller that is responsible
% for mixing the individuals from different subpopulations and producing new
% subpopulations. These new subpopulations are pushed again by the controller into
% the message queue, creating a loop. The cycle stops when the controller receives
% a subpopulation containing a candidate solution that satisfies a particular
% condition, or a maximum number of messages were received.

@@

\item The Controller process is responsible for keeping track of the progress of 
the search. It pulls current subpopulations from the Output queue, inspects the metadata 
and if an optimal solution has been found or the maximum number of iteration has been
reached it signals the stop of the execution. Otherwise, it passes the stream of messages 
to a migration process, where subpopulations are mixed with each other. New populations 
are generated from this migration, and they are again pushed to the Input queue to continue 
in a loop. The Controller is also responsible of logging the metadata recived with the
messages.  
\end{enumerate}

This reactive architecture has the following advantages:

An important aspect of proposal is the decoupling of the population and 
the population-based algorithm. It is common that in a classic island-based algorithm
each island is executed in a separate processing node, i.e. in a virtual machine, cpu,
or thread. In this case we can have just one processing node, or many nodes, each 
running a different search strategy, or using its own parameters.

Also, the reactive controller, gives designers more control over the milti-population algorithm.
In this process, designers can dynamicly change the number of populations, population parameters,
and migration details on-the-fly.

Another advantege is that algorithm designers have many options for implementing this simple architecure. 
The same basic components can be implmented as a single multithreaded program, or 
as a highly scalable serverless cloud application. Most modern languages include 
constructs for asynchronous programming using queues or channels, for a multi-threaded
execution.

On the other hand, there are several caveats as well:

It is more costly to move entire subpopulations as messages than only passing certain individuals 
from a process to another. Designers must conmsider this cost when working with large
individuals, and if possible send sub-populations using several messages, use compression or 
adjust the size of subpopulations. Pool-based algorithms also suffer from this drawback, but 
web based immplementations have been working with continous optimization problems of 1000 dimensions.
But, in other cases this is not a viable approach.

@@

\subsection{Comparasion} 
\label{edm}

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{classicisland}
    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
    \label{fig:kafkEO}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{pool_island}
    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
    \label{fig:kafkEO}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{population_pool}
    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
    \label{fig:kafkEO}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{population_message}
    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
    \label{fig:kafkEO}
\end{figure}
